---
title: "Basic Inferential data analysis"
author: "Tuan Vu"
date: "Saturday, October 25, 2014"
output: pdf_document
---

*This is the second part of Statistical Inference Project.*

**Requirements**
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 
Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.
Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions. 

*Let's do some analysis*

**Load the data**
```{r}
library(ggplot2)
data(ToothGrowth)
```

*1. Summary of the data*

**Take a look at first few rows**
```{r}
head(ToothGrowth)
```

**Type of Supplements and Dosages**
```{r}
table(ToothGrowth$supp) # Supplement
table(ToothGrowth$dose) # Dosage
```

**Summary of ToothGrowth**
```{r}
summary(ToothGrowth)
```

**Correlation in ToothGrowth data**
```{r}
cor(ToothGrowth[sapply(ToothGrowth, is.numeric)])
```

We can see that the length of the tooth and the dosage have a high correlation

**Plotting**
```{r}
p1 <- ggplot(ToothGrowth, aes(x =factor(dose), y = len, fill = factor(dose)))
p1 + geom_boxplot() + xlab("Dosage") + ylab("Length")
```

Again, the boxplot shows more clearly that as the dosage increases, the length of the tooth increases.

*2. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose*

**The 95% Confidence Intervals of Tooth Grow**
```{r}
Interval <- (mean(ToothGrowth$len) + c(-1, 1) * qnorm(0.975) * sd(ToothGrowth$len)/sqrt(length(ToothGrowth$len)))
Interval
```

**T-Test by length and Supplements**
```{r}
t.test(len~supp, data=ToothGrowth)
```

This T-Test shows that we cannot reject the Null hypothesis.

**T-Test of Dosage Level**

Subset Data
```{r}
dos1 <- subset(ToothGrowth, dose == 0.5)
dos2 <- subset(ToothGrowth, dose == 1)
dos3 <- subset(ToothGrowth, dose == 2)
```

T-Test
```{r}
t.test(len~supp, data=dos1) # Small Dosage = 0.5
t.test(len~supp, data=dos2) # Medium Dosage = 1
t.test(len~supp, data=dos3) # Big Dosage = 2
```

- For Small and Medium dosage, their p-values are small, so we can reject the null hypothesis
- For Big Dosage, we cannot reject the null hypothesis

*3. State your conclusions and the assumptions needed for your conclusions*

- The Tooth Length Growth is not controlled by Supplements.
- For dosage level 0.5 and 1, the Orange Juice has a higher effect on the length of tooth of Guinea Pigs than the Vitamin C.
- For dosage level 2, there are so such difference between Orange Juice and Vitamin C
